---
title: "c4: Covariate Connectivity Combined Clustering"
output: github_document
bibliography: inst/REFERENCES.bib
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

[![License: MIT](https://img.shields.io/badge/License-MIT-green.svg)](LICENSE)

## Covariate Connectivity Combined Clustering ($\text{C}^4$)

The \texttt{c4} package implements the Covariate Connectivity Combined Clustering ($\text{C}^4$) method, which integrates both network topology and node-level covariates to improve community detection in weighted networks.

This R package provides functions to:

- simulate weighted networks under stochastic block models with user-defined block structures and weight distributions;

- perform $\text{C}^4$ for joint network-covariate community detection [@hu2025covariate];

- apply covariate-assisted spectral clustering as a competing method [@binkiewicz2017covariate].

## Installation

You can install $\texttt{c4}$ from github with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("zyhu888/c4")
```

## References
